---
title: "01_r_overview"
output: html_document
---

# docs/01_r_overview.Rmd

# 1. What is R?

R is a programming language and environment designed for statistical computing and data analysis.

It is widely used in:

-   Data analysis
-   Statistics
-   Machine learning
-   Data visualization
-   Research and academia

Key features of R:

-   Powerful statistical functions
-   Rich visualization libraries
-   Large ecosystem of packages
-   Open source

# 2. Running R Code

R code can be executed in several ways:

-   R console
-   RStudio
-   R scripts (`.R`)
-   R Markdown (`.Rmd`)

Example:

```{r (Hello R)}
print("Hello R")
```

When this document is rendered, the code will run and display the result.

# 3. Basic Arithmetic

R can be used like a calculator.

```{r calculator}
1 + 1
10 - 3
5 * 4 
20 / 5
```

# 4. Creating Variables

In R, variables are usually assigned using `<-`.

```{r (using `<-`)}
x <- 10
y <- 20

x
y
```

# 5. Simple Calculation

```{r Simple Calculation}
result <- x + y
result
```

# 6. Data Types

R has several basic data types.

```{r data types}
num <- 10
text <- "Hello"
flag <- TRUE

class(num)
class(text)
class(flag)
```

# 7. Getting Help

R has built-in help documentation.

```{r eval=FALSE}
help(mean)
?mean
```

You can also run examples from documentation.

```{r eval=FALSE}
example(mean)
```

# 8. Summary

In this chapter we learned:

-   What R is
-   How to run R code
-   Basic arithmetic
-   Variables
-   Data types
-   How to access help

Next chapter:

**02_basic_syntax.R**
