---
title: "02 Basic Syntax"
author: "Toshi"
output: html_document
---

# 1. Introduction

In this chapter, we learn the basic syntax of R.

Topics covered:

-   Comments
-   Assignment
-   Arithmetic operations
-   Basic object types
-   Checking object types
-   Listing and removing objects

# 2. Comments

Comments in R start with `#`.

They are used to explain the code and are ignored during execution.

```{r}
# This is a comment
print("Hello from R")
```

# 3. Assignment

In R, values are usually assigned with `<-`.

```{r}
x <- 10
y <- 3

x
y
```

# 4. Arithmetic Operations

R can perform basic arithmetic operations.

```{r}
x + y
x - y
x * y
x / y
x ^ y
```

# 5. Character and Logical Values

R can store text and logical values.

```{r}
name <- "Toshi"
is_learning_r <- TRUE

name
is_learning_r
```

# 6. Checking Object Types

Use `class()` to check the type of an object.

```{r}
class(x)
class(name)
class(is_learning_r)
```

# 7. Using print()

`print()` displays the value of an object clearly.

```{r}
print(x)
print(name)
print(is_learning_r)
```

# 8. Listing Objects

Use `ls()` to see the objects currently stored in the environment.

```{r}
ls()
```

# 9. Removing Objects

Use `rm()` to remove objects you no longer need.

```{r}
temp_value <- 100
ls()

rm(temp_value)
ls()
```

# 10. Summary

In this chapter, we learned:

-   how to write comments
-   how to assign values
-   how to do arithmetic
-   how to create basic objects
-   how to check object types
-   how to list and remove objects

Next chapter:

**03_vectors.Rmd**
