---
title: "03 Vectors"
author: "Toshi"
output: html_document
---

# 1. Introduction

Vectors are one of the most important data structures in R.

A vector is a collection of values of the same type.

Examples:

-   numeric vector
-   character vector
-   logical vector

# 2. Creating Vectors

Use `c()` to combine values into a vector.

```{r}
numbers <- c(10, 20, 30, 40, 50)
fruits <- c("apple", "banana", "orange")
flags <- c(TRUE, FALSE, TRUE)

numbers
fruits
flags
```

# 3. Checking the Type of a Vector

Use `class()` to check the type.

```{r}
class(numbers)
class(fruits)
class(flags)
```

# 4. Length of a Vector

Use `length()` to count the number of elements.

```{r}
length(numbers)
length(fruits)
length(flags)
```

# 5. Basic Statistics for Numeric Vectors

You can apply summary functions to numeric vectors.

```{r}
sum(numbers)
mean(numbers)
max(numbers)
min(numbers)
```

# 6. Vectorized Operations

R can apply operations to all elements of a vector at once.

```{r}
numbers + 5
numbers * 2
numbers / 10
```

# 7. Accessing Elements

Use square brackets `[]` to access elements.

```{r}
numbers[1]
numbers[3]
fruits[2]
```

# 8. Accessing Multiple Elements

You can extract multiple elements by passing a vector of positions.

```{r}
numbers[c(1, 3, 5)]
fruits[c(1, 2)]
```

# 9. Excluding Elements

Use negative indices to remove elements.

```{r}
numbers[-1]
numbers[-c(2, 4)]
```

# 10. Creating Sequences

R provides convenient ways to create sequential vectors.

```{r}
1:10
seq(0, 20, by = 5)
```

# 11. Repeating Values

Use `rep()` to repeat values.

```{r}
rep(1, 5)
rep(c("A", "B"), 3)
```

# 12. Logical Filtering

Logical conditions can be used to filter values.

```{r}
numbers > 25
numbers[numbers > 25]
```

# 13. Combining Vectors

You can combine vectors using `c()`.

```{r}
more_numbers <- c(60, 70)
all_numbers <- c(numbers, more_numbers)

all_numbers
```

# 14. Summary

In this chapter, we learned:

-   how to create vectors
-   how to check vector types
-   how to count elements
-   how to calculate summaries
-   how to extract values
-   how to filter values
-   how to create sequences and repeated values

Next chapter:

**04_data_frame.Rmd**
