---
title: "04_data_frame"
output: html_document
---

# 1. What is a Data Frame?

A data frame is one of the most important data structures in R.

It looks like a table:

-   rows = observations
-   columns = variables

A data frame can store different types of data in different columns.

# 2. Creating a Data Frame

Use `data.frame()` to create a table-like object.

```{r}
students <- data.frame(
  id = c(1, 2, 3, 4, 5),
  name = c("Alice", "Bob", "Carol", "David", "Emma"),
  age = c(20, 21, 19, 22, 20),
  score = c(85, 90, 78, 92, 88)
)

students
```

# 3. Checking the Structure

You can inspect the structure of a data frame using several functions.

```{r}
class(students)
str(students)
dim(students)
nrow(students)
ncol(students)
names(students)
```

# 4. Accessing Columns

You can access columns by name or position.

```{r}
students$name
students$score
students[, "age"]
students[, 4]
```

# 5. Accessing Rows

You can access rows using row numbers.

```{r}
students[1, ]
students[3, ]
students[1:3, ]
```

# 6. Accessing Specific Rows and Columns

You can choose both rows and columns at the same time.

```{r}
students[1:3, c("name", "score")]
students[, c("name", "age")]
```

# 7. Filtering Rows

You can filter rows by conditions.

```{r}
students[students$score >= 85, ]
students[students$age == 20, ]
```

# 8. Adding a New Column

You can add a new column using `$`.

```{r}
students$passed <- students$score >= 80
students
```

# 9. Creating a Calculated Column

A new column can be created from existing columns.

```{r}
students$score_plus_5 <- students$score + 5
students
```

# 10. Summary Statistics

You can calculate summary statistics for numeric columns.

```{r}
mean(students$score)
max(students$score)
min(students$score)
summary(students)
```

# 11. Sorting Rows

Use `order()` to sort a data frame.

```{r}
students[order(students$score), ]
students[order(-students$score), ]
```

# 12. Summary

In this chapter, we learned:

-   what a data frame is
-   how to create a data frame
-   how to inspect its structure
-   how to access rows and columns
-   how to filter rows
-   how to add new columns
-   how to summarize and sort data

Next chapter:

**05_basic_statistics.Rmd**
