---
title: "05_basic_statistics"
output: html_document
---

# 1. Introduction

In this chapter, we learn basic statistical functions in R.

These functions help us summarize and understand numeric data.

# 2. Prepare Sample Data

```{r}
scores <- c(85, 90, 78, 92, 88, 76, 95, 89, 84, 91)

scores
```

# 3. Basic Summary Values

R provides simple functions to calculate key summary values.

```{r}
sum(scores)
mean(scores)
median(scores)
min(scores)
max(scores)
range(scores)
```

# 4. Measure of Spread

We can also measure how spread out the data is.

```{r}
var(scores)
sd(scores)
```

# 5. Quantiles

Quantiles divide data into parts.

```{r}
quantile(scores)
quantile(scores, probs = c(0.25, 0.5, 0.75))
```

# 6. Sorting Values

Sorting helps us inspect the order of values.

```{r}
sort(scores)
sort(scores, decreasing = TRUE)
```

# 7. Counting Values

Use `length()` to count the number of elements.

```{r}
length(scores)
```

# 8. Summary Function

`summary()` gives a quick overview of numeric data.

```{r}
summary(scores)
```

# 9. Basic Statistics on a Data Frame Column

Statistical functions are often applied to one column of a data frame.

```{r}
students <- data.frame(
  id = c(1, 2, 3, 4, 5),
  name = c("Alice", "Bob", "Carol", "David", "Emma"),
  score = c(85, 90, 78, 92, 88)
)

students

mean(students$score)
summary(students$score)
```

# 10. Logical Conditions

We can also check which values satisfy a condition.

```{r}
scores > 85
scores[scores > 85]
```

# 11. Missing Values

Sometimes data contains missing values (`NA`).

```{r}
scores_with_na <- c(85, 90, NA, 92, 88)

scores_with_na
mean(scores_with_na)
mean(scores_with_na, na.rm = TRUE)
summary(scores_with_na)
```

`na.rm = TRUE` means "remove missing values before calculation".

# 12. Summary

In this chapter, we learned:

-   how to calculate mean, median, min, max
-   how to measure spread with variance and standard deviation
-   how to calculate quantiles
-   how to use summary()
-   how to handle missing values with `na.rm = TRUE`

Next chapter:

**06_visualization.Rmd**
