---
title: "06_visualization"
output: html_document
---

# 1. Introduction

In this chapter, we learn how to visualize data in R.

Visualization helps us understand patterns and trends in data.

# 2. Prepare Sample Data

```{r}
scores <- c(85, 90, 78, 92, 88, 76, 95, 89, 84, 91)
scores
```

# 3. Scatter Plot

Use `plot()` to create a scatter plot.

```{r}
x <- 1:length(scores)

plot(x, scores)
```

# 4. Customize Plot

You can add titles and labels.

```{r}
plot(
  x, scores,
  main = "Score Distribution",
  xlab = "Index",
  ylab = "Score",
  pch = 19
)
```

# 5. Line Plot

Use `type = "l"` to draw lines.

```{r}
plot(
  x, scores,
  type = "l",
  main = "Score Trend",
  xlab = "Index",
  ylab = "Score"
)
```

# 6. Histogram

Use `hist()` to see the distribution of data.

```{r}
hist(
  scores,
  main = "Histogram of Scores",
  xlab = "Score",
  col = "lightblue"
)
```

# 7. Boxplot

Use `boxplot()` to see summary statistics visually.

```{r}
boxplot(
  scores,
  main = "Boxplot of Scores",
  col = "lightgreen"
)
```

# 8. Barplot

Use `barplot()` for categorical data.

```{r}
counts <- table(c("A", "B", "A", "C", "B", "A"))

barplot(
  counts,
  main = "Category Count",
  col = "orange"
)
```

# 9. Plot from Data Frame

You can plot values directly from a data frame.

```{r}
students <- data.frame(
  id = 1:5,
  score = c(85, 90, 78, 92, 88)
)

plot(
  students$id,
  students$score,
  main = "Student Scores",
  xlab = "ID",
  ylab = "Score",
  pch = 19
)
```

# 10. Summary

In this chapter, we learned:

-   how to create scatter plots
-   how to customize plots
-   how to create line plots
-   how to create histograms
-   how to create boxplots
-   how to create barplots

Next chapter:

**07_data_analysis.Rmd**
