---
title: "07_real_data_analysis"
output: html_document
---

# 1. Introduction

In this chapter, we practice a simple real data analysis workflow in R.

The workflow is:

1.  read a CSV file
2.  inspect the data
3.  summarize the data
4.  visualize the data

# 2. Read CSV

We use `read.csv()` to load a CSV file.

```{r}
sales_data <- read.csv("data/sample_sales.csv", stringsAsFactors = FALSE)

sales_data
```

# 3. Inspect Data

Before analysis, it is important to understand the structure of the data.

```{r}
head(sales_data)
str(sales_data)
dim(sales_data)
names(sales_data)
summary(sales_data)
```

# 4. Basic Summaries

We can calculate basic statistics for numeric columns.

```{r}
sum(sales_data$sales)
mean(sales_data$sales)
max(sales_data$sales)
min(sales_data$sales)
```

# 5. Filter Rows

We can extract rows that match specific conditions.

```{r}
sales_data[sales_data$region == "Tokyo", ]
sales_data[sales_data$category == "Beverage", ]
```

# 6. Aggregate by Category

We can summarize values by category using `aggregate()`.

```{r}
aggregate(sales ~ category, data = sales_data, sum)
aggregate(quantity ~ category, data = sales_data, sum)
```

# 7. Aggregate by Region

We can also summarize sales by region.

```{r}
aggregate(sales ~ region, data = sales_data, sum)
```

# 8. Count by Product

Use `table()` to count occurrences of values.

```{r}
table(sales_data$product)
```

# 9. Visualize Data

We can create simple charts from real data.

```{r}
hist(
  sales_data$sales,
  main = "Histogram of Sales",
  xlab = "Sales",
  col = "lightblue"
)

boxplot(
  sales_data$sales,
  main = "Boxplot of Sales",
  col = "lightgreen"
)

barplot(
  table(sales_data$category),
  main = "Count by Category",
  col = "orange"
)
```

# 10. Visualize Aggregated Data

We can visualize summarized data as well.

```{r}
sales_by_region <- aggregate(sales ~ region, data = sales_data, sum)

barplot(
  sales_by_region$sales,
  names.arg = sales_by_region$region,
  main = "Total Sales by Region",
  col = "skyblue"
)
```

# 11. Summary

In this chapter, we learned:

-   how to read a CSV file
-   how to inspect data
-   how to summarize numeric columns
-   how to filter rows
-   how to aggregate values by group
-   how to create basic charts from real data

Next chapter:

**08_data_cleaning.Rmd**
