---
title: "08_data_cleaning"
output: html_document
---

# 1. Introduction

In real-world analysis, raw data is often messy.

Typical problems include:

-   missing values
-   extra spaces
-   inconsistent text
-   wrong data types

In this chapter, we learn how to clean such data in R.

# 2. Create Sample Dirty Data

```{r}
sales_dirty <- data.frame(
  id = c(1, 2, 3, 4, 5),
  product = c(" Coffee", "Tea ", "Bread", NA, "Cake"),
  sales = c("450", "300", "280", "520", NA),
  quantity = c(3, 2, NA, 2, 4),
  region = c("tokyo", "Osaka", "TOKYO", "Nagoya", "osaka"),
  stringsAsFactors = FALSE
)

sales_dirty
```

# 3. Inspect the Data

We first check the structure and summary.

```{r}
str(sales_dirty)
summary(sales_dirty)
```

# 4. Check Missing Values

Use `is.na()` to detect missing values.

```{r}
is.na(sales_dirty)
colSums(is.na(sales_dirty))
```

`colSums(is.na(...))` shows how many missing values each column has.

# 5. Remove Rows with Missing Values

Use `na.omit()` to remove rows containing missing values.

```{r}
sales_no_na <- na.omit(sales_dirty)
sales_no_na
```

# 6. Trim Extra Spaces

Use `trimws()` to remove extra spaces from text.

```{r}
sales_dirty$product <- trimws(sales_dirty$product)
sales_dirty
```

# 7. Convert Text to Numeric

Sometimes numeric values are stored as text.

```{r}
sales_dirty$sales <- as.numeric(sales_dirty$sales)
sales_dirty
```

# 8. Standardize Text Values

Text values may have inconsistent capitalization.

```{r}
sales_dirty$region <- tolower(sales_dirty$region)
sales_dirty
```

# 9. Replace Values

After standardizing case, we can replace values with consistent labels.

```{r}
sales_dirty$region[sales_dirty$region == "tokyo"] <- "Tokyo"
sales_dirty$region[sales_dirty$region == "osaka"] <- "Osaka"
sales_dirty$region[sales_dirty$region == "nagoya"] <- "Nagoya"

sales_dirty
```

# 10. Fill Missing Numeric Values

We can replace missing numeric values when needed.

```{r}
sales_dirty$quantity[is.na(sales_dirty$quantity)] <- 0
sales_dirty
```

# 11. Create a New Flag Column

We can create a column to mark missing values.

```{r}
sales_dirty$is_missing_sales <- is.na(sales_dirty$sales)
sales_dirty
```

# 12. Create a Category by Condition

Use `ifelse()` to create categories.

```{r}
sales_dirty$sales_level <- ifelse(
  sales_dirty$sales >= 400,
  "High",
  "Low"
)

sales_dirty
```

# 13. Recheck the Cleaned Data

After cleaning, inspect the data again.

```{r}
str(sales_dirty)
summary(sales_dirty)
```

# 14. Summary

In this chapter, we learned:

-   how to detect missing values
-   how to remove rows with missing values
-   how to trim spaces
-   how to convert text to numeric
-   how to standardize text values
-   how to replace values
-   how to create new columns with conditions

Next chapter:

**09_next_steps.Rmd**
