---
title: "09_ggplot2_basics"
output: html_document
---

# 1. Introduction

In this chapter, we learn basic data visualization using `ggplot2`.

`ggplot2` is one of the most popular visualization packages in R.

It allows us to build charts by combining layers.

# 2. Load ggplot2

If `ggplot2` is not installed, install it first.

```{r eval=FALSE}
install.packages("ggplot2")
```

Then load the package.

```{r}
library(ggplot2)
```

# 3. Create Sample Data

```{r}
students <- data.frame(
  id = 1:8,
  name = c("Alice", "Bob", "Carol", "David", "Emma", "Frank", "Grace", "Henry"),
  score = c(85, 90, 78, 92, 88, 76, 95, 89),
  study_hours = c(3, 4, 2, 5, 4, 2, 6, 5),
  class = c("A", "A", "B", "B", "A", "B", "A", "B")
)

students
```

# 4. Basic Scatter Plot

A scatter plot shows the relationship between two numeric variables.

```{r}
ggplot(students, aes(x = study_hours, y = score)) +
  geom_point()
```

# 5. Scatter Plot with Color

We can use `color` to separate groups.

```{r}
ggplot(students, aes(x = study_hours, y = score, color = class)) +
  geom_point()
```

# 6. Add Title and Labels

Use `labs()` to add a title and axis labels.

```{r}
ggplot(students, aes(x = study_hours, y = score, color = class)) +
  geom_point() +
  labs(
    title = "Study Hours and Scores",
    x = "Study Hours",
    y = "Score",
    color = "Class"
  )
```

# 7. Bar Chart

A bar chart is useful for counting categories.

```{r}
ggplot(students, aes(x = class)) +
  geom_bar()
```

# 8. Boxplot

A boxplot helps us compare distributions between groups.

```{r}
ggplot(students, aes(x = class, y = score)) +
  geom_boxplot()
```

# 9. Histogram

A histogram shows the distribution of numeric data.

```{r}
ggplot(students, aes(x = score)) +
  geom_histogram(binwidth = 5)
```

# 10. Add a Smooth Line

We can add a regression line using `geom_smooth()`.

```{r}
ggplot(students, aes(x = study_hours, y = score)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

# 11. ggplot2 Basic Structure

The basic structure of ggplot2 is:

``` r
ggplot(data, aes(x = column1, y = column2)) +
  geom_xxx()
```

Main parts:

-   `ggplot()` : specify data
-   `aes()` : define mapping
-   `geom_*()` : choose chart type
-   `labs()` : add labels

# 12. Summary

In this chapter, we learned:

-   how to load ggplot2
-   how to create scatter plots
-   how to use color
-   how to add titles and labels
-   how to create bar charts
-   how to create boxplots
-   how to create histograms
-   how to add a smooth line
