---
title: "10_functions"
output: html_document
---

# 1. Introduction

In this chapter, we learn how to create reusable functions in R.

Functions help us avoid writing the same code again and again.

# 2. Prepare Sample Data

```{r}
scores <- c(85, 90, 78, 92, 88)

scores
```

# 3. Create a Simple Function

Use `function()` to create a function.

```{r}
say_hello <- function() {
  print("Hello from a function")
}

say_hello()
```

# 4. Function with One Argument

A function can receive input values called arguments.

```{r}
calculate_average <- function(x) {
  mean(x)
}

calculate_average(scores)
```

# 5. Return Values

Use `return()` to explicitly return a result.

```{r}
calculate_total <- function(x) {
  total <- sum(x)
  return(total)
}

calculate_total(scores)
```

# 6. Function with Multiple Arguments

A function can use multiple arguments.

```{r}
calculate_score_rate <- function(score, max_score) {
  rate <- score / max_score
  return(rate)
}

calculate_score_rate(85, 100)
```

# 7. Default Arguments

Default arguments are used when no value is provided.

```{r}
add_bonus <- function(score, bonus = 5) {
  score + bonus
}

add_bonus(80)
add_bonus(80, bonus = 10)
```

# 8. Use Functions with a Data Frame

Functions can also be applied to columns in a data frame.

```{r}
students <- data.frame(
  id = c(1, 2, 3, 4, 5),
  name = c("Alice", "Bob", "Carol", "David", "Emma"),
  score = c(85, 90, 78, 92, 88)
)

students

calculate_average(students$score)
calculate_total(students$score)
```

# 9. Create a Reusable Summary Function

We can combine multiple calculations into one function.

```{r}
summarize_scores <- function(x) {
  result <- data.frame(
    average = mean(x),
    total = sum(x),
    minimum = min(x),
    maximum = max(x)
  )
  
  return(result)
}

summarize_scores(scores)
summarize_scores(students$score)
```

# 10. Reuse Functions with Different Data

Once a function is created, it can be reused with different data.

```{r}
scores_class_a <- c(85, 90, 78, 92, 88)
scores_class_b <- c(72, 80, 91, 86, 79)

calculate_average(scores_class_a)
calculate_average(scores_class_b)

summarize_scores(scores_class_a)
summarize_scores(scores_class_b)
```

# 11. Summary

In this chapter, we learned:

- how to create a function
- how to use arguments
- how to return values
- how to use default arguments
- how to create reusable analysis functions
- how to apply functions to vectors and data frame columns

Next chapter:

**11_program_flow.Rmd**