---
title: "11_program_flow"
output: html_document
---

# 1. Introduction

In this chapter, we learn how to control the flow of a program using conditions and loops.

Program flow allows us to make decisions and repeat tasks automatically.

# 2. Prepare Sample Data

```{r}
scores <- c(85, 90, 78, 92, 88)

scores
```

# 3. if Statement

Use `if` to execute code only when a condition is TRUE.

```{r}
score <- 85

if (score >= 80) {
  print("Pass")
}
```

# 4. if ... else

Use `if ... else` to choose between two actions.

```{r}
score <- 75

if (score >= 80) {
  print("Pass")
} else {
  print("Fail")
}
```

# 5. ifelse()

`ifelse()` evaluates multiple values at once.

```{r}
ifelse(scores >= 80, "Pass", "Fail")
```

# 6. for Loop

A `for` loop repeats code for each element.

```{r}
for (score in scores) {
  print(score)
}
```

# 7. while Loop

A `while` loop repeats while a condition remains TRUE.

```{r}
count <- 1

while (count <= 5) {
  print(count)
  count <- count + 1
}
```

# 8. break

Use `break` to stop a loop.

```{r}
for (score in scores) {

  if (score == 92) {
    break
  }

  print(score)

}
```

# 9. next

Use `next` to skip the current iteration.

```{r}
for (score in scores) {

  if (score < 80) {
    next
  }

  print(score)

}
```

# 10. Apply Program Flow to Data Analysis

Extract scores greater than or equal to 90.

```{r}
high_scores <- c()

for (score in scores) {

  if (score >= 90) {
    high_scores <- c(high_scores, score)
  }

}

high_scores
```

# 11. Compare with the Average

Find scores above the average.

```{r}
average_score <- mean(scores)

above_average <- c()

for (score in scores) {

  if (score > average_score) {
    above_average <- c(above_average, score)
  }

}

average_score
above_average
```

# 12. Summary

In this chapter, we learned:

-   how to use `if`
-   how to use `if ... else`
-   how to use `ifelse()`
-   how to repeat processing with `for`
-   how to use `while`
-   how to stop a loop with `break`
-   how to skip an iteration with `next`
-   how to apply program flow to simple data analysis

Next chapter:

**12_matrix_list_factor.Rmd**
