---
title: "12_matrix_list_factor"
output: html_document
---

# 1. Introduction

In this chapter, we learn three additional data structures in R:

- Matrix
- List
- Factor

These structures are useful for organizing different kinds of data in analysis.

# 2. Create a Matrix

A matrix stores values in rows and columns.

Unlike a data frame, all values in a matrix are usually the same data type.

```{r}
scores_matrix <- matrix(
  c(85, 90, 78, 92, 88, 81),
  nrow = 2,
  ncol = 3,
  byrow = TRUE
)

scores_matrix
```

# 3. Check Matrix Structure

We can inspect the dimensions of a matrix.

```{r}
class(scores_matrix)
dim(scores_matrix)
nrow(scores_matrix)
ncol(scores_matrix)
```

# 4. Access Matrix Elements

Use `[row, column]` to access values.

```{r}
scores_matrix[1, 1]
scores_matrix[2, 3]
scores_matrix[1, ]
scores_matrix[, 2]
```

# 5. Basic Matrix Calculations

Functions can be applied to the entire matrix, rows, or columns.

```{r}
sum(scores_matrix)
mean(scores_matrix)
rowMeans(scores_matrix)
colMeans(scores_matrix)
```

# 6. Create a List

A list can store different types of objects together.

```{r}
student <- list(
  name = "Alice",
  age = 20,
  scores = c(85, 90, 88)
)

student
```

# 7. Access List Elements

List elements can be accessed by name or position.

```{r}
student$name
student$age
student$scores

student[["name"]]
student[[3]]
```

# 8. Modify a List

We can add new elements or update existing values.

```{r}
student$passed <- TRUE
student$age <- 21

student
```

# 9. Create a Factor

A factor represents categorical data.

```{r}
grades <- c("A", "B", "A", "C", "B", "A")

grade_factor <- factor(grades)

grade_factor
class(grade_factor)
levels(grade_factor)
```

# 10. Count Factor Levels

Use `table()` to count each category.

```{r}
table(grade_factor)
```

# 11. Create an Ordered Factor

Factors can also represent categories with a meaningful order.

```{r}
satisfaction <- c("Low", "High", "Medium", "High", "Low")

satisfaction_factor <- factor(
  satisfaction,
  levels = c("Low", "Medium", "High"),
  ordered = TRUE
)

satisfaction_factor
levels(satisfaction_factor)
```

# 12. Convert Between Data Structures

R allows us to convert data structures when needed.

```{r}
matrix_as_data_frame <- as.data.frame(scores_matrix)
matrix_as_data_frame

as.vector(scores_matrix)
```

# 13. Use Multiple Data Structures Together

Lists are useful for storing different analysis results in one object.

```{r}
analysis_result <- list(
  student = student,
  scores = scores_matrix,
  grades = grade_factor
)

analysis_result
```

# 14. Summary

In this chapter, we learned:

- how to create and inspect matrices
- how to access rows and columns in a matrix
- how to calculate row and column averages
- how to create and modify lists
- how to represent categorical data with factors
- how to create ordered factors
- how to convert between data structures
- how to combine different objects in a list

Next chapter:

**13_string_processing.Rmd**
