---
title: "13_string_processing"
output: html_document
---

# 1. Introduction

Real-world data is rarely clean.

Character data may contain:

- Extra spaces
- Inconsistent uppercase and lowercase letters
- Different naming conventions
- Unwanted characters
- Values that do not follow the expected format

Before analyzing data, we often need to validate and standardize these values.

String processing is therefore an important part of **data cleaning and data management**.

In this chapter, we learn basic R functions for inspecting, cleaning, searching, and transforming character data.

# 2. Prepare Sample Text Data

First, prepare character data that contains inconsistent formatting.

```{r}
products <- c(" Coffee ", "GREEN TEA", "bread", "Chocolate Cake ")

products
```

# 3. Count Characters

Use `nchar()` to count the number of characters in each value.

This can also help identify unexpected values during data validation.

```{r}
nchar(products)
```

# 4. Combine Strings

Use `paste()` and `paste0()` to combine character values.

```{r}
product <- "Coffee"
region <- "Tokyo"

paste(product, region)
paste(product, region, sep = " - ")
paste0(product, "_", region)
```

# 5. Change Letter Case

Inconsistent uppercase and lowercase values can cause problems when grouping or comparing data.

Use `toupper()` and `tolower()` to standardize letter case.

```{r}
toupper(products)
tolower(products)
```

# 6. Remove Extra Spaces

Extra spaces are common in imported data.

Use `trimws()` to remove spaces from the beginning and end of character values.

```{r}
trimmed_products <- trimws(products)

trimmed_products
```

# 7. Extract Part of a String

Use `substr()` to extract part of a character value.

```{r}
substr("Chocolate Cake", 1, 9)
substr(trimmed_products[1], 1, 3)
```

# 8. Split Strings

Some datasets store multiple pieces of information in one character value.

Use `strsplit()` to split the value.

```{r}
product_code <- "FOOD-COFFEE-001"

strsplit(product_code, "-")
```

# 9. Replace Text

Use `gsub()` to replace text patterns.

```{r}
product_names <- c("Green_Tea", "Chocolate_Cake", "Coffee")

gsub("_", " ", product_names)
```

# 10. Search Text

Use `grep()` to find matching values and `grepl()` to check whether each value matches a pattern.

```{r}
products_clean <- trimws(products)

grep("Tea", products_clean, value = TRUE)
grepl("Cake", products_clean)
```

# 11. Validate Character Data

Validation checks whether values follow expected rules.

For example, region values can be compared with a list of valid values.

```{r}
regions <- c("Tokyo", "Osaka", "Nagoya", "TOKYO", "Unknown")

valid_regions <- c("Tokyo", "Osaka", "Nagoya")

regions %in% valid_regions
```

Notice that `"TOKYO"` is not equal to `"Tokyo"`.

This is why standardization is important before validation.

# 12. Clean Character Data in a Data Frame

Now apply string processing to data that resembles a real dataset.

```{r}
sales_data <- data.frame(
  product = c(" Coffee ", "GREEN TEA", "bread", "Chocolate Cake "),
  region = c("tokyo", "OSAKA", "Nagoya", "TOKYO"),
  stringsAsFactors = FALSE
)

sales_data
```

Remove extra spaces and standardize letter case.

```{r}
sales_data$product <- trimws(sales_data$product)
sales_data$product <- tolower(sales_data$product)

sales_data$region <- trimws(sales_data$region)
sales_data$region <- tolower(sales_data$region)

sales_data
```

# 13. Standardize Text Values

Use `tools::toTitleCase()` to make the cleaned values easier to read.

```{r}
sales_data$product <- tools::toTitleCase(sales_data$product)
sales_data$region <- tools::toTitleCase(sales_data$region)

sales_data
```

The values now follow a consistent format.

# 14. Validate Cleaned Data

After cleaning, validate the region values again.

```{r}
valid_regions <- c("Tokyo", "Osaka", "Nagoya")

sales_data$is_valid_region <- sales_data$region %in% valid_regions

sales_data
```

This creates a validation flag that can be used to identify invalid records.

# 15. Create Labels from Character Data

Clean character values can also be combined to create readable labels.

```{r}
sales_data$label <- paste(
  sales_data$product,
  sales_data$region,
  sep = " - "
)

sales_data
```

# 16. From Raw Data to Analysis-Ready Data

The basic workflow in this chapter is:

``` text
Raw Character Data
        ↓
Inspect
        ↓
Clean
        ↓
Standardize
        ↓
Validate
        ↓
Analysis-Ready Data
```

String processing is not only about changing text.

It is an important technique for improving **data quality** before analysis.

# 17. Summary

In this chapter, we learned:

- why character data needs validation and cleaning
- how to count characters with `nchar()`
- how to combine strings with `paste()` and `paste0()`
- how to standardize letter case
- how to remove extra spaces
- how to extract and split strings
- how to replace and search text
- how to validate values against expected data
- how to clean character columns in a data frame
- how string processing supports data quality and data management

Next chapter:

**14_date_time.Rmd**
