---
title: "14_date_time"
output: html_document
---

# 1. Introduction

Real-world datasets often contain dates and timestamps.

Examples include:

- Order dates
- Registration dates
- Update timestamps
- Event times
- Monthly and yearly records

Date and time data is important because many analyses depend on **when** something happened.

However, values such as `"2026-08-16"` may initially be stored as character data.

Before time-based analysis, we often need to:

- Convert text into date/time objects
- Format dates consistently
- Calculate differences between dates
- Extract year, month, or weekday
- Filter data by time periods

In this chapter, we learn the basics of handling date and time data in R.

# 2. Check the Current Date and Time

Use `Sys.Date()` and `Sys.time()` to check the current date and time.

```{r}
Sys.Date()
Sys.time()
```

# 3. Create Date Objects

Use `as.Date()` to create a Date object.

```{r}
order_date <- as.Date("2026-08-16")

order_date
class(order_date)
```

# 4. Convert Character Data to Date

Date values imported from CSV files may be stored as character data.

```{r}
date_text <- c("2026-08-01", "2026-08-05", "2026-08-10")

dates <- as.Date(date_text)

dates
class(dates)
```

This conversion is important before performing date calculations.

# 5. Format Dates

Use `format()` to extract or change the displayed format of a date.

```{r}
format(order_date, "%Y")
format(order_date, "%m")
format(order_date, "%d")
format(order_date, "%Y/%m/%d")
format(order_date, "%A")
```

Common format codes include:

- `%Y` : four-digit year
- `%m` : month
- `%d` : day
- `%A` : weekday name

# 6. Add and Subtract Days

Date objects can be used in arithmetic operations.

```{r}
order_date + 7
order_date - 7
```

This is useful for creating deadlines, windows, or comparison periods.

# 7. Calculate Date Differences

We can calculate the difference between two dates.

```{r}
start_date <- as.Date("2026-08-01")
end_date <- as.Date("2026-08-16")

end_date - start_date
difftime(end_date, start_date, units = "days")
```

# 8. Create a Sequence of Dates

Use `seq.Date()` to generate consecutive dates.

```{r}
date_sequence <- seq.Date(
  from = as.Date("2026-08-01"),
  to = as.Date("2026-08-07"),
  by = "day"
)

date_sequence
```

# 9. Create Time Objects

Use `as.POSIXct()` for date and time values.

```{r}
event_time <- as.POSIXct("2026-08-16 14:30:00")

event_time
class(event_time)
```

`Date` is useful when only the calendar date matters.

`POSIXct` is useful when hours, minutes, and seconds are also important.

# 10. Format Date and Time

The `format()` function can also be used with timestamps.

```{r}
format(event_time, "%Y-%m-%d")
format(event_time, "%H:%M:%S")
format(event_time, "%Y-%m-%d %H:%M")
```

# 11. Convert Date Columns in a Data Frame

Real-world data often contains date columns stored as character values.

```{r}
sales_data <- data.frame(
  order_id = c(1, 2, 3, 4),
  order_date = c("2026-08-01", "2026-08-03", "2026-08-05", "2026-08-08"),
  sales = c(450, 300, 520, 280),
  stringsAsFactors = FALSE
)

sales_data
str(sales_data)
```

Convert the date column before analysis.

```{r}
sales_data$order_date <- as.Date(sales_data$order_date)

sales_data
str(sales_data)
```

# 12. Extract Time-Based Information

Once the column is stored as Date, we can create useful analysis fields.

```{r}
sales_data$year <- format(sales_data$order_date, "%Y")
sales_data$month <- format(sales_data$order_date, "%m")
sales_data$weekday <- format(sales_data$order_date, "%A")

sales_data
```

These fields can later be used for grouping and comparison.

### About Weekday Names

The `%A` format code returns the full weekday name for a Date value.

For example:

```{r}
format(as.Date("2026-08-01"), "%A")
```

Common date format codes include:

- %Y : four-digit year
- %m : month
- %d : day
- %A : full weekday name
- %a : abbreviated weekday name

The weekday name returned by %A depends on the locale of the R environment.

For example, the same date may be displayed as Saturday in an English locale or 土曜日 in a Japanese locale.

# 13. Calculate Elapsed Days

We can calculate how many days have passed since each event.

```{r}
reference_date <- as.Date("2026-08-10")

sales_data$days_since_order <- as.numeric(
  reference_date - sales_data$order_date
)

sales_data
```

# 14. Filter Data by Date

Dates can be used directly in filter conditions.

```{r}
sales_data[
  sales_data$order_date >= as.Date("2026-08-05"),
]
```

# 15. Summarize Sales by Date

Date columns can also be used as grouping fields.

```{r}
aggregate(
  sales ~ order_date,
  data = sales_data,
  sum
)
```

# 16. From Character Data to Time-Based Analysis

The basic workflow in this chapter is:

``` text
Character Data
      ↓
Convert to Date / Time
      ↓
Format
      ↓
Extract Time Information
      ↓
Calculate Differences
      ↓
Filter and Summarize
      ↓
Time-Based Analysis
```

Date and time handling is therefore an important part of preparing real-world data for analysis.

# 17. Summary

In this chapter, we learned:

- how to check the current date and time
- how to create Date objects
- how to convert character values to dates
- how to format dates
- how to add and subtract days
- how to calculate date differences
- how to create date sequences
- how to create POSIXct time objects
- how to convert date columns in a data frame
- how to extract year, month, and weekday
- how to filter and summarize data by date

Next chapter:

**15_data_manipulation_dplyr.Rmd**
