---
title: "Chapter 15: Data Manipulation with dplyr"
output: html_document
---

# Chapter 15: Data Manipulation with dplyr

## Purpose

In previous chapters, we manipulated data frames using base R.

In real data analysis, we often repeat operations such as selecting columns, filtering rows, creating calculated columns, sorting data, grouping records, and calculating summaries.

The `dplyr` package provides a clear and readable way to perform these operations.

The goal of this chapter is not to memorize every `dplyr` function. Instead, we will learn how to combine a small set of important functions into a practical data analysis workflow.

**Select / Filter → Transform → Group → Summarize → Sort**

## 1. Install and Load dplyr

`dplyr` is an external R package.

Install it once if necessary:

```{r eval=FALSE}
install.packages("dplyr")
```

Then load it:

```{r}
library(dplyr)
```

## 2. Prepare Sample Data

```{r}
sales_data <- data.frame(
  order_id = 1:10,
  region = c("Tokyo","Osaka","Tokyo","Fukuoka","Osaka",
             "Tokyo","Fukuoka","Tokyo","Osaka","Tokyo"),
  category = c("Books","Food","Food","Books","Books",
               "Electronics","Food","Books","Electronics","Food"),
  product = c("Book A","Coffee","Tea","Book B","Book C",
              "Mouse","Snack","Book D","Keyboard","Juice"),
  price = c(1200,500,300,1500,1000,2500,400,1800,4500,600),
  quantity = c(2,3,5,1,2,1,4,2,1,3)
)
sales_data
```

## 3. Select Columns

`select()` chooses the columns that we want to keep.

```{r}
selected_data <- sales_data |>
  select(order_id, region, category, price)
selected_data
```

A similar operation with base R is:

```{r}
sales_data[c("order_id", "region", "category", "price")]
```

Both approaches are valid. `dplyr` provides syntax designed to make a sequence of data manipulation steps easy to read.

## 4. Filter Rows

`filter()` keeps rows that satisfy a condition.

```{r}
tokyo_sales <- sales_data |>
  filter(region == "Tokyo")
tokyo_sales
```

A similar operation in base R is:

```{r}
sales_data[sales_data$region == "Tokyo", ]
```

## 5. Sort Rows

`arrange()` sorts rows.

```{r}
sorted_sales <- sales_data |>
  arrange(desc(price))
sorted_sales
```

`desc()` means descending order. Without `desc()`, values are sorted in ascending order.

## 6. Create Calculated Columns

`mutate()` creates or modifies columns.

**Total Sales = Price × Quantity**

```{r}
sales_with_total <- sales_data |>
  mutate(total_sales = price * quantity)
sales_with_total
```

## 7. Summarize Data

`summarise()` calculates summary values.

```{r}
sales_summary <- sales_data |>
  summarise(
    total_quantity = sum(quantity),
    average_price = mean(price),
    maximum_price = max(price)
  )
sales_summary
```

## 8. Group Data

`group_by()` defines groups for later operations.

```{r}
grouped_sales <- sales_data |>
  group_by(category)
grouped_sales
```

`group_by()` by itself does not calculate a summary. It defines the groups that later operations, especially `summarise()`, should use.

## 9. Group and Summarize

Combining `group_by()` and `summarise()` is one of the most important patterns in `dplyr`.

```{r}
category_summary <- sales_data |>
  mutate(total_sales = price * quantity) |>
  group_by(category) |>
  summarise(
    orders = n(),
    quantity = sum(quantity),
    sales = sum(total_sales),
    .groups = "drop"
  )
category_summary
```

Here, `n()` counts rows in each group. `.groups = "drop"` returns the result without keeping the grouping structure.

## 10. Use Multiple Conditions

```{r}
filtered_sales <- sales_data |>
  filter(
    region %in% c("Tokyo", "Osaka"),
    price >= 1000
  )
filtered_sales
```

Conditions separated by commas inside `filter()` work like AND.

```{r}
books_or_food <- sales_data |>
  filter(category == "Books" | category == "Food")
books_or_food
```

| Operator | Meaning |
|---|---|
| `&` | AND |
| `|` | OR |
| `!` | NOT |
| `%in%` | Value is included in a set |

## 11. Chain Operations with the Pipe

The native R pipe `|>` passes the result of one operation to the next.

```{r}
tokyo_summary <- sales_data |>
  filter(region == "Tokyo") |>
  mutate(total_sales = price * quantity) |>
  group_by(category) |>
  summarise(
    sales = sum(total_sales),
    .groups = "drop"
  )
tokyo_summary
```

Read it from top to bottom:

1. Start with `sales_data`.
2. Keep Tokyo records.
3. Calculate total sales.
4. Group by category.
5. Calculate sales for each category.

### `|>` and `%>%`

You may also see `%>%` in many R programs and tutorials. `%>%` became popular through the `magrittr` package and is widely used with `dplyr`.

This manual uses the native R pipe `|>` for its main examples. Existing R code may use either form.

## 12. Build a Simple Analysis Workflow

Suppose we want to answer:

> Which product category generates the most sales in Tokyo?

```{r}
analysis_result <- sales_data |>
  filter(region == "Tokyo") |>
  mutate(total_sales = price * quantity) |>
  group_by(category) |>
  summarise(
    orders = n(),
    quantity = sum(quantity),
    sales = sum(total_sales),
    .groups = "drop"
  ) |>
  arrange(desc(sales))
analysis_result
```

The workflow is:

```text
Raw sales data
      |
      v
Filter Tokyo records
      |
      v
Calculate total sales
      |
      v
Group by category
      |
      v
Summarize each group
      |
      v
Sort by sales
      |
      v
Analysis result
```

## Summary

| Function | Purpose |
|---|---|
| `select()` | Select columns |
| `filter()` | Filter rows |
| `arrange()` | Sort rows |
| `mutate()` | Create or modify columns |
| `summarise()` | Calculate summaries |
| `group_by()` | Define groups |
| `n()` | Count rows in a group |

The main idea is:

**Data manipulation is usually a sequence of small operations.**

By connecting those operations into a readable workflow, we can transform raw data into analysis-ready results.

In the next chapter, we will extend data manipulation to working with multiple datasets through joins and reshaping.
