---
title: "Chapter 16: Data Joining and Reshaping"
output: html_document
---

# Chapter 16: Data Joining and Reshaping

## Purpose

Real-world data is often stored across multiple tables.

For example:

- sales records may contain a product ID,
- a product master may contain product names and prices,
- a customer master may contain customer information.

Before analysis, these datasets often need to be combined.

We may also need to change the shape of a dataset so that it is easier to analyze or visualize.

In this chapter, we will learn how to:

- join multiple datasets,
- check unmatched records,
- combine rows,
- convert between wide and long formats.

The main workflow is:

**Join → Validate → Transform → Summarize → Reshape**

---

## 1. Install and Load Packages

This chapter uses `dplyr` and `tidyr`.

Install them once if necessary:

```{r eval=FALSE}
install.packages("dplyr")
install.packages("tidyr")
```

Then load them:

```{r}
library(dplyr)
library(tidyr)
```

---

## 2. Prepare Sample Datasets

We will use a sales table and a product master.

```{r}
sales_data <- data.frame(
  order_id = 1:6,
  product_id = c("P001", "P002", "P001", "P003", "P004", "P999"),
  quantity = c(2, 1, 3, 2, 1, 4)
)

product_data <- data.frame(
  product_id = c("P001", "P002", "P003", "P004"),
  product_name = c("Book A", "Mouse", "Coffee", "Keyboard"),
  category = c("Books", "Electronics", "Food", "Electronics"),
  price = c(1200, 2500, 500, 4500)
)

sales_data
product_data
```

Notice that `P999` exists in `sales_data`, but not in `product_data`.

We will use this unmatched value later.

---

## 3. Understand Keys

A join needs a column that connects two datasets.

Here, the common column is:

```text
product_id
```

In a master table such as `product_data`, this key should normally identify each product uniquely.

```{r}
product_data$product_id
duplicated(product_data$product_id)
```

If duplicate keys exist in a master table, a join may create unexpected duplicate rows.

This is why checking keys is an important validation step.

---

## 4. Join Datasets with left_join()

`left_join()` keeps all rows from the left dataset.

```{r}
sales_with_products <- sales_data |>
  left_join(product_data, by = "product_id")

sales_with_products
```

Because `P999` does not exist in the product master, its product information becomes `NA`.

This behavior is useful when we want to keep the original transaction rows and inspect missing master data.

---

## 5. Join Matching Rows with inner_join()

`inner_join()` keeps only rows with matching keys in both datasets.

```{r}
matched_sales <- sales_data |>
  inner_join(product_data, by = "product_id")

matched_sales
```

The row with `P999` is removed because no matching product exists.

---

## 6. Compare left_join() and inner_join()

A simple way to understand the difference is to compare row counts.

```{r}
nrow(sales_data)
nrow(sales_with_products)
nrow(matched_sales)
```

In general:

| Join | Result |
|---|---|
| `left_join()` | Keep all rows from the left table |
| `inner_join()` | Keep only matching rows |

Choosing a join type depends on the purpose of the analysis.

---

## 7. Join Using Different Column Names

Sometimes two datasets use different names for the same key.

For example:

```{r}
sales_alt <- data.frame(
  order_id = 1:3,
  item_code = c("P001", "P002", "P003"),
  quantity = c(1, 2, 1)
)

sales_alt
```

The sales table uses `item_code`, while the product table uses `product_id`.

We can connect them with `join_by()`.

```{r}
joined_alt <- sales_alt |>
  left_join(
    product_data,
    by = join_by(item_code == product_id)
  )

joined_alt
```

This tells `dplyr` that `item_code` and `product_id` represent the same join key.

---

## 8. Check Unmatched Records

Before continuing an analysis, it is often useful to check whether any records failed to match.

`anti_join()` returns rows from the left dataset that have no match in the right dataset.

```{r}
unmatched_sales <- sales_data |>
  anti_join(product_data, by = "product_id")

unmatched_sales
```

This is useful for data validation.

For example, unmatched records may indicate:

- missing master data,
- incorrect IDs,
- old codes,
- data entry errors.

A join should not be treated only as a technical operation.  
The result should also be validated.

---

## 9. Combine Rows

Sometimes datasets have the same columns but contain records from different periods or sources.

For example:

```{r}
sales_january <- data.frame(
  order_id = 101:103,
  product_id = c("P001", "P002", "P003"),
  quantity = c(1, 2, 1)
)

sales_february <- data.frame(
  order_id = 104:106,
  product_id = c("P001", "P004", "P002"),
  quantity = c(2, 1, 3)
)
```

`bind_rows()` stacks them vertically.

```{r}
combined_sales <- bind_rows(
  sales_january,
  sales_february
)

combined_sales
```

This is useful when combining monthly, yearly, or partitioned files.

---

## 10. Understand Wide and Long Data

Data can be stored in different shapes.

A wide dataset may look like this:

```{r}
monthly_sales_wide <- data.frame(
  product = c("Book", "Mouse"),
  Jan = c(100, 80),
  Feb = c(120, 90),
  Mar = c(150, 110)
)

monthly_sales_wide
```

Conceptually:

```text
product  Jan  Feb  Mar
Book     100  120  150
Mouse     80   90  110
```

A long version contains one row per product and month:

```text
product  month  sales
Book     Jan    100
Book     Feb    120
Book     Mar    150
Mouse    Jan     80
...
```

Long format is often easier to use for grouping, filtering, and visualization.

---

## 11. Convert Wide Data to Long Format

`pivot_longer()` converts multiple columns into rows.

```{r}
monthly_sales_long <- monthly_sales_wide |>
  pivot_longer(
    cols = Jan:Mar,
    names_to = "month",
    values_to = "sales"
  )

monthly_sales_long
```

Here:

- `cols = Jan:Mar` selects the columns to reshape.
- `names_to = "month"` stores the original column names.
- `values_to = "sales"` stores the values.

---

## 12. Convert Long Data to Wide Format

`pivot_wider()` performs the opposite transformation.

```{r}
monthly_sales_wide_again <- monthly_sales_long |>
  pivot_wider(
    names_from = month,
    values_from = sales
  )

monthly_sales_wide_again
```

This converts the long dataset back into a wide structure.

---

## 13. Build a Multi-Table Analysis Workflow

Now we can combine the techniques from Chapters 15 and 16.

Suppose we want to calculate sales by category.

```{r}
analysis_result <- sales_data |>
  left_join(product_data, by = "product_id") |>
  filter(!is.na(product_name)) |>
  mutate(total_sales = price * quantity) |>
  group_by(category) |>
  summarise(
    orders = n(),
    quantity = sum(quantity),
    sales = sum(total_sales),
    .groups = "drop"
  ) |>
  arrange(desc(sales))

analysis_result
```

The process is:

```text
Sales data
    |
    v
Join product master
    |
    v
Check / remove unmatched records
    |
    v
Calculate total sales
    |
    v
Group by category
    |
    v
Summarize
    |
    v
Sort results
```

This is a common pattern in real data analysis.

---

## Summary

In this chapter, we learned:

| Function | Purpose |
|---|---|
| `left_join()` | Keep all rows from the left dataset |
| `inner_join()` | Keep only matching rows |
| `anti_join()` | Find unmatched rows |
| `bind_rows()` | Combine datasets vertically |
| `pivot_longer()` | Convert wide data to long format |
| `pivot_wider()` | Convert long data to wide format |

The important idea is:

**Joining data is not only about combining tables. It also requires validation.**

After joining datasets, we should check unmatched records, duplicate keys, and the shape of the result before continuing the analysis.

In the next chapter, we will build more advanced visualizations with `ggplot2`.
