---
title: "Chapter 17: Advanced ggplot2"
output: html_document
---

# Chapter 17: Advanced ggplot2

## Purpose

Chapter 09 introduced the basic structure of `ggplot2`.

In this chapter, we move from simply creating charts to using visualization as part of the data analysis process.

The goal is not to learn every visual option available in `ggplot2`.

Instead, we will learn how to make analysis results:

- easier to read,
- easier to compare,
- easier to interpret, and
- easier to communicate.

The workflow is:

**Analyze → Visualize → Compare → Customize → Communicate**

------------------------------------------------------------------------

## 1. Install and Load Packages

This chapter uses `dplyr` for data manipulation and `ggplot2` for visualization.

```{r eval=FALSE}
install.packages("dplyr")
install.packages("ggplot2")
```

Load the packages:

```{r}
library(dplyr)
library(ggplot2)
```

------------------------------------------------------------------------

## 2. Prepare Sample Sales Data

We will use sales data containing three regions and three product categories.

```{r}
sales_data <- data.frame(
  region = c(
    "Tokyo", "Tokyo", "Tokyo",
    "Osaka", "Osaka", "Osaka",
    "Fukuoka", "Fukuoka", "Fukuoka"
  ),
  category = rep(c("Books", "Electronics", "Food"), 3),
  sales = c(
    7200, 9800, 5600,
    6100, 11200, 4900,
    4800, 7600, 5900
  )
)

sales_data
```

------------------------------------------------------------------------

## 3. Review the ggplot2 Structure

The basic structure remains the same as Chapter 09.

```{r}
ggplot(
  sales_data,
  aes(x = category, y = sales)
) +
  geom_col()
```

The main components are:

``` text
Data
  +
Aesthetic mapping: aes()
  +
Geometry: geom_*()
```

`aes()` defines how variables are mapped to visual properties such as x position, y position, color, and fill.

`geom_*()` defines how those values are displayed.

------------------------------------------------------------------------

## 4. Visualize Summarized Data

A visualization often starts after data has been transformed or summarized.

First, summarize sales by category with `dplyr`.

```{r}
category_sales <- sales_data |>
  group_by(category) |>
  summarise(
    sales = sum(sales),
    .groups = "drop"
  )

category_sales
```

Then visualize the result.

```{r}
ggplot(
  category_sales,
  aes(x = category, y = sales)
) +
  geom_col()
```

This creates an important workflow:

**Raw Data → Transform / Summarize → Visualize**

`dplyr` prepares the data, while `ggplot2` communicates the result.

------------------------------------------------------------------------

## 5. Customize Titles and Labels

`labs()` adds information that helps readers understand a chart.

```{r}
ggplot(
  category_sales,
  aes(x = category, y = sales)
) +
  geom_col() +
  labs(
    title = "Sales by Category",
    subtitle = "Total sales across all regions",
    x = "Category",
    y = "Sales"
  )
```

A chart should provide enough context for the reader to understand what is being shown.

------------------------------------------------------------------------

## 6. Customize Themes

Themes control the non-data appearance of a plot.

For example:

```{r}
ggplot(
  category_sales,
  aes(x = category, y = sales)
) +
  geom_col() +
  theme_minimal()
```

Another built-in theme is:

```{r}
ggplot(
  category_sales,
  aes(x = category, y = sales)
) +
  geom_col() +
  theme_classic()
```

Individual elements can also be adjusted with `theme()`.

```{r}
ggplot(
  category_sales,
  aes(x = category, y = sales)
) +
  geom_col() +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(title = "Sales by Category")
```

The purpose of customization is not decoration.

It should improve readability and communication.

------------------------------------------------------------------------

## 7. Control Categories and Axes

Categories are not always displayed in the most useful order.

`reorder()` can order categories by a numeric value.

```{r}
ggplot(
  category_sales,
  aes(
    x = reorder(category, sales),
    y = sales
  )
) +
  geom_col()
```

`coord_flip()` switches the axes.

```{r}
ggplot(
  category_sales,
  aes(
    x = reorder(category, sales),
    y = sales
  )
) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Sales by Category",
    x = "Category",
    y = "Sales"
  ) +
  theme_minimal()
```

Horizontal bar charts can be useful when category names are long or when ranking categories.

------------------------------------------------------------------------

## 8. Use Color and Fill Effectively

Visual properties can represent additional variables.

For a bar chart, `fill` controls the inside of the bars.

```{r}
ggplot(
  sales_data,
  aes(
    x = category,
    y = sales,
    fill = region
  )
) +
  geom_col(position = "dodge") +
  labs(
    title = "Sales by Category and Region",
    x = "Category",
    y = "Sales",
    fill = "Region"
  ) +
  theme_minimal()
```

In `ggplot2`:

- `color` usually controls lines and outlines.
- `fill` usually controls filled areas such as bars.

Manual scales such as `scale_fill_manual()` can be used when a specific color definition is required.

Colors should help distinguish information rather than simply decorate the chart.

------------------------------------------------------------------------

## 9. Compare Groups with Facets

Facets create separate panels for groups in the data.

`facet_wrap()` is useful when one grouping variable should be displayed across multiple panels.

```{r}
ggplot(
  sales_data,
  aes(x = category, y = sales)
) +
  geom_col() +
  facet_wrap(~ region) +
  labs(
    title = "Sales by Category for Each Region",
    x = "Category",
    y = "Sales"
  ) +
  theme_minimal()
```

`facet_grid()` arranges panels using a row-and-column structure.

```{r}
ggplot(
  sales_data,
  aes(x = category, y = sales)
) +
  geom_col() +
  facet_grid(region ~ .) +
  theme_minimal()
```

Facets are useful when we want to compare the same measurement across groups without placing everything in one chart.

------------------------------------------------------------------------

## 10. Add Reference and Trend Information

A reference line can provide context.

First calculate the average sales value.

```{r}
average_sales <- mean(sales_data$sales)
```

Then add it with `geom_hline()`.

```{r}
ggplot(
  sales_data,
  aes(x = category, y = sales)
) +
  geom_point() +
  geom_hline(
    yintercept = average_sales,
    linetype = "dashed"
  ) +
  facet_wrap(~ region) +
  labs(
    title = "Sales Compared with Overall Average",
    x = "Category",
    y = "Sales"
  ) +
  theme_minimal()
```

For numeric relationships or trends, `geom_smooth()` can add a fitted line.

```{r}
trend_data <- data.frame(
  month = 1:8,
  sales = c(100, 120, 115, 145, 150, 165, 170, 195)
)

ggplot(
  trend_data,
  aes(x = month, y = sales)
) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Sales Trend",
    x = "Month",
    y = "Sales"
  ) +
  theme_minimal()
```

A trend line can help us see a pattern, but the visual pattern alone does not prove a statistical relationship.

Statistical analysis will be introduced in the next chapter.

------------------------------------------------------------------------

## 11. Add Text Annotations

`geom_text()` can display values associated with observations.

```{r}
ggplot(
  category_sales,
  aes(
    x = reorder(category, sales),
    y = sales
  )
) +
  geom_col() +
  geom_text(
    aes(label = sales),
    hjust = -0.1
  ) +
  coord_flip() +
  labs(
    title = "Sales by Category",
    x = "Category",
    y = "Sales"
  ) +
  theme_minimal()
```

`annotate()` can add explanatory text at a specific position.

```{r}
ggplot(
  trend_data,
  aes(x = month, y = sales)
) +
  geom_line() +
  geom_point() +
  annotate(
    "text",
    x = 8,
    y = 195,
    label = "Highest sales",
    vjust = -1
  ) +
  theme_minimal()
```

Annotations should highlight useful information rather than overload the chart.

------------------------------------------------------------------------

## 12. Save Plots

A completed plot can be stored in an object.

```{r}
final_plot <- ggplot(
  category_sales,
  aes(
    x = reorder(category, sales),
    y = sales
  )
) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Sales by Category",
    x = "Category",
    y = "Sales"
  ) +
  theme_minimal()

final_plot
```

`ggsave()` saves a plot to a file.

```{r eval=FALSE}
ggsave(
  filename = "sales_by_category.png",
  plot = final_plot,
  width = 8,
  height = 5
)
```

This is useful when a visualization needs to be included in a report, presentation, or other output.

------------------------------------------------------------------------

## 13. Build an Analysis Visualization

Finally, combine data manipulation and visualization.

First summarize the data.

```{r}
region_category_sales <- sales_data |>
  group_by(region, category) |>
  summarise(
    sales = sum(sales),
    .groups = "drop"
  )
```

Then create the visualization.

```{r}
analysis_plot <- ggplot(
  region_category_sales,
  aes(
    x = reorder(category, sales),
    y = sales
  )
) +
  geom_col() +
  coord_flip() +
  facet_wrap(~ region) +
  labs(
    title = "Category Sales by Region",
    subtitle = "Comparison of sales patterns across regions",
    x = "Category",
    y = "Sales"
  ) +
  theme_minimal()

analysis_plot
```

The complete workflow is:

``` text
Raw Data
    |
    v
Transform / Summarize
    |
    v
Visualize
    |
    v
Compare
    |
    v
Customize
    |
    v
Communicate
```

Visualization is not separate from data analysis.

It is one of the ways we explore results, compare patterns, and communicate findings.

------------------------------------------------------------------------

## Summary

In this chapter, we learned how to move beyond basic `ggplot2` charts.

| Function        | Purpose                         |
|-----------------|---------------------------------|
| `labs()`        | Add titles and labels           |
| `theme_*()`     | Apply a built-in theme          |
| `theme()`       | Customize plot elements         |
| `reorder()`     | Reorder categories              |
| `coord_flip()`  | Flip axes                       |
| `facet_wrap()`  | Create panels by group          |
| `facet_grid()`  | Arrange panels in a grid        |
| `geom_hline()`  | Add a horizontal reference line |
| `geom_smooth()` | Add a fitted trend line         |
| `geom_text()`   | Add labels from data            |
| `annotate()`    | Add explanatory annotations     |
| `ggsave()`      | Save a plot                     |

The main idea is:

**A good visualization should make an analysis result easier to understand.**

The goal is not to add as many visual elements as possible. Each customization should have a reason.

In the next chapter, we will use statistical analysis to examine relationships and differences in data more formally.
