---
title: "Chapter 18: Statistical Analysis"
output: html_document
---

# Chapter 18: Statistical Analysis

## Purpose

Chapter 05 introduced descriptive statistics such as the mean, median, variance, and standard deviation.

Chapter 17 showed how visualization can reveal patterns and differences in data.

In this chapter, we take the next step and use basic statistical methods to examine those patterns more formally.

The goal is not to study statistical theory in depth. Instead, we will learn how R can help us answer questions such as:

- Are two numeric variables related?
- Can one variable be used to estimate another?
- Is there a difference between groups?
- Are two categorical variables associated?

The workflow is:

**Question → Explore → Visualize → Test / Model → Interpret**

---

## 1. Prepare Sample Data

First, create a small dataset containing advertising and sales values.

```{r}
sales_data <- data.frame(
  advertising = c(10, 12, 15, 18, 20, 23, 25, 28, 30, 35),
  sales = c(100, 115, 130, 145, 155, 170, 180, 195, 205, 230)
)

sales_data
```

We will also prepare score data for three groups.

```{r}
score_data <- data.frame(
  group = c(
    rep("A", 8),
    rep("B", 8),
    rep("C", 8)
  ),
  score = c(
    72, 75, 78, 74, 80, 77, 76, 79,
    82, 85, 81, 88, 84, 86, 83, 87,
    68, 70, 72, 69, 74, 71, 73, 70
  )
)

score_data
```

---

## 2. Review Descriptive Statistics

Before performing a statistical test or building a model, we should understand the data itself.

```{r}
mean(sales_data$sales)
median(sales_data$sales)
sd(sales_data$sales)
summary(sales_data$sales)
```

We can also compare group averages.

```{r}
aggregate(
  score ~ group,
  data = score_data,
  FUN = mean
)
```

Descriptive statistics provide context for the analysis that follows.

---

## 3. Understand Correlation

Correlation measures the strength and direction of a relationship between two numeric variables.

```{r}
correlation <- cor(
  sales_data$advertising,
  sales_data$sales
)

correlation
```

A correlation coefficient is generally between `-1` and `1`.

```text
Close to  1 : strong positive relationship
Close to  0 : weak linear relationship
Close to -1 : strong negative relationship
```

A positive correlation means that larger values of one variable tend to occur with larger values of the other.

However:

**Correlation does not necessarily mean causation.**

Even when two variables are strongly correlated, this alone does not prove that one variable causes the other.

---

## 4. Visualize a Relationship

Before interpreting a correlation value, it is useful to inspect the relationship visually.

```{r}
library(ggplot2)

ggplot(
  sales_data,
  aes(x = advertising, y = sales)
) +
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE
  ) +
  labs(
    title = "Advertising and Sales",
    x = "Advertising",
    y = "Sales"
  ) +
  theme_minimal()
```

The scatter plot shows the individual observations.

The fitted line helps us see the overall linear trend.

Statistical values and visualizations should be interpreted together.

---

## 5. Build a Simple Linear Regression Model

`lm()` creates a linear model.

```{r}
model <- lm(
  sales ~ advertising,
  data = sales_data
)

model
```

The formula:

```r
sales ~ advertising
```

can be read as:

> Explain or estimate `sales` using `advertising`.

In this example:

- `sales` is the response variable.
- `advertising` is the explanatory variable.

The purpose here is to understand the basic structure of regression in R rather than the full mathematical theory.

---

## 6. Inspect a Regression Model

`summary()` provides more information about the fitted model.

```{r}
summary(model)
```

The output contains several pieces of statistical information, including:

- estimated coefficients,
- residual information,
- R-squared,
- p-values.

At this stage, focus on the overall idea:

**The model describes a relationship found in the observed data.**

A model should still be interpreted together with the original data, visualization, and analysis question.

---

## 7. Make Predictions with a Model

A fitted model can be used to estimate values for new observations.

```{r}
new_data <- data.frame(
  advertising = c(16, 22, 32)
)

predicted_sales <- predict(
  model,
  newdata = new_data
)

predicted_sales
```

We can combine the input and predicted values.

```{r}
prediction_result <- data.frame(
  advertising = new_data$advertising,
  predicted_sales = predicted_sales
)

prediction_result
```

Prediction does not mean certainty.

The result is an estimate based on the fitted model and the data used to create it.

---

## 8. Compare Two Groups

Suppose we want to compare the scores of groups A and B.

First, select the two groups.

```{r}
two_group_data <- subset(
  score_data,
  group %in% c("A", "B")
)
```

Then perform a t-test.

```{r}
t_test_result <- t.test(
  score ~ group,
  data = two_group_data
)

t_test_result
```

A t-test can be used to examine whether two groups differ in their mean values.

The test result includes a p-value, which we will discuss later in this chapter.

---

## 9. Compare Multiple Groups

When we have more than two groups, analysis of variance (ANOVA) can be used to examine whether group means differ.

```{r}
anova_model <- aov(
  score ~ group,
  data = score_data
)

summary(anova_model)
```

In this example, we compare groups A, B, and C.

A significant ANOVA result indicates evidence that not all group means are equal.

It does not, by itself, tell us exactly which groups differ from each other.

---

## 10. Examine Categorical Relationships

Statistical analysis is not limited to numeric variables.

Suppose we have device type and purchase status.

```{r}
survey_data <- data.frame(
  device = c(
    "PC", "PC", "PC", "PC", "PC",
    "Mobile", "Mobile", "Mobile", "Mobile", "Mobile",
    "Tablet", "Tablet", "Tablet", "Tablet", "Tablet"
  ),
  purchased = c(
    "Yes", "Yes", "No", "Yes", "No",
    "Yes", "No", "No", "Yes", "No",
    "No", "No", "Yes", "No", "No"
  )
)
```

Create a contingency table with `table()`.

```{r}
survey_table <- table(
  survey_data$device,
  survey_data$purchased
)

survey_table
```

Then perform a chi-squared test.

```{r}
chi_square_result <- chisq.test(survey_table)

chi_square_result
```

This test examines whether two categorical variables appear to be associated.

For small datasets, R may display a warning because the expected counts are too small for the usual chi-squared approximation. This is an important reminder that statistical tests have assumptions and should not be applied mechanically.

---

## 11. Understand p-values

Many statistical tests return a p-value.

For example:

```{r}
t_test_result$p.value
```

A commonly used reference level is:

```text
p-value < 0.05
```

When the p-value is below a chosen significance level such as `0.05`, the observed result is considered statistically significant under the assumptions of the test.

However, this should **not** be interpreted as:

```text
p < 0.05 = important
p >= 0.05 = unimportant
```

A p-value does not directly tell us:

- how large an effect is,
- whether the result is practically important,
- whether one variable causes another.

It is one piece of evidence in the analysis.

---

## 12. Statistical and Practical Meaning

Suppose a statistical test finds a difference between groups.

We should still examine the actual values.

```{r}
group_summary <- aggregate(
  score ~ group,
  data = score_data,
  FUN = mean
)

group_summary
```

Consider both questions:

1. Is there statistical evidence of a difference?
2. Is the size of the difference meaningful for the real problem?

Statistical significance and practical importance are not the same thing.

This is why interpretation is an essential part of statistical analysis.

---

## 13. Build a Statistical Analysis Workflow

Now combine the ideas from this chapter.

### Question

> Is advertising associated with sales?

### Step 1: Explore the data

```{r}
summary(sales_data)
```

### Step 2: Measure correlation

```{r}
cor(
  sales_data$advertising,
  sales_data$sales
)
```

### Step 3: Visualize the relationship

```{r}
analysis_plot <- ggplot(
  sales_data,
  aes(x = advertising, y = sales)
) +
  geom_point() +
  geom_smooth(
    method = "lm",
    se = FALSE
  ) +
  labs(
    title = "Relationship between Advertising and Sales",
    x = "Advertising",
    y = "Sales"
  ) +
  theme_minimal()

analysis_plot
```

### Step 4: Build and inspect a model

```{r}
analysis_model <- lm(
  sales ~ advertising,
  data = sales_data
)

summary(analysis_model)
```

### Step 5: Interpret

Do not interpret only one number.

Consider:

- the original question,
- descriptive statistics,
- the visualization,
- the strength of the relationship,
- the statistical output,
- the practical meaning.

The complete workflow is:

```text
Question
   |
   v
Explore Data
   |
   v
Visualize
   |
   v
Test / Model
   |
   v
Interpret
```

---

## Summary

In this chapter, we learned several basic statistical tools available in R.

| Function | Purpose |
|---|---|
| `cor()` | Measure correlation |
| `lm()` | Build a linear regression model |
| `summary()` | Inspect statistical output |
| `predict()` | Make predictions from a model |
| `t.test()` | Compare two group means |
| `aov()` | Compare multiple group means |
| `table()` | Create a contingency table |
| `chisq.test()` | Examine categorical association |

The main idea is:

**Statistical analysis starts with a question and ends with interpretation.**

A function result should not be treated as an answer by itself.

Explore the data, visualize it, choose an appropriate method, and interpret the result in the context of the original question.

In the next chapter, we will learn how to import data from and export analysis results to external files.
