---
title: "Chapter 19: Import and Export"
output: html_document
---

# Chapter 19: Import and Export

## Purpose

So far, many examples in this manual have created data directly inside R.

In real data analysis, however, data usually comes from external sources such as CSV, text, or Excel files.

After analysis, the results may also need to be used outside R.

This chapter introduces the basic input/output workflow:

**External Data → Import → Inspect → Analyze → Export**

---

## 1. Understand Data Input and Output

Two basic operations connect R with external files.

**Import** means reading data from an external file into R.

**Export** means writing data or analysis results from R to an external file.

For example:

```text
CSV File
   |
   v
Import into R
   |
   v
Analyze
   |
   v
Export Result
   |
   v
CSV / Excel / Text
```

Import and export make it possible to use R as one part of a larger data workflow.

---

## 2. Check the Working Directory

R uses a working directory as the starting point for relative file paths.

```{r}
getwd()
```

This manual uses paths such as:

```text
data/sample_sales_19.csv
output/analysis_result.csv
```

These are **relative paths**.

For the examples in this chapter, run the script from the repository root so that the `data` and `output` directories can be found correctly.

---

## 3. Import CSV Files

CSV is one of the most common formats for exchanging tabular data.

Chapter 07 already used `read.csv()`. Here, we review it as part of the complete input/output workflow.

```{r}
sales_data <- read.csv(
  "data/sample_sales_19_19.csv"
)

sales_data
```

`read.csv()` reads the file and returns a data frame.

---

## 4. Inspect Imported Data

Importing a file successfully does not automatically mean that the data is ready for analysis.

Always inspect the imported data.

```{r}
head(sales_data)
str(sales_data)
summary(sales_data)
dim(sales_data)
```

These functions help answer questions such as:

- Were the expected rows imported?
- Are the column names correct?
- Are numeric columns numeric?
- Are there missing values?
- Does the dataset have the expected dimensions?

A useful rule is:

**Import → Inspect before Analyze**

---

## 5. Import Text Files

Delimited text files can be imported with `read.table()`.

The sample file in this chapter uses tab-separated values.

```{r}
text_data <- read.table(
  "data/sample_sales_19.txt",
  header = TRUE,
  sep = "\t"
)

text_data
```

Important arguments include:

| Argument | Purpose |
|---|---|
| `header` | Whether the first row contains column names |
| `sep` | Character used to separate fields |

Different text files may use tabs, commas, spaces, or other delimiters.

---

## 6. Import Excel Files

Excel files are common in business environments.

The `readxl` package can read Excel workbooks.

Install it once if necessary:

```{r eval=FALSE}
install.packages("readxl")
```

Load the package:

```{r}
library(readxl)
```

Then read an Excel file:

```{r}
excel_data <- read_excel(
  "data/sample_sales_19.xlsx"
)

excel_data
```

`read_excel()` can read both `.xls` and `.xlsx` files.

The result can then be inspected and analyzed like other tabular data in R.

---

## 7. Handle File Paths Safely

A script should not depend unnecessarily on one person's computer.

For example, this absolute path may work only on one machine:

```r
"E:/R/R-Basic-Manual/data/sample_sales_19.csv"
```

Inside a project, prefer a relative path:

```{r}
csv_path <- "data/sample_sales_19.csv"

sales_from_path <- read.csv(csv_path)

head(sales_from_path)
```

This makes the project easier to move to another folder or computer.

A good project structure might look like:

```text
R-Basic-Manual/
├─ data/
│  └─ sample_sales_19.csv
├─ scripts/
├─ docs/
└─ output/
```

Relative paths improve portability and reproducibility.

---

## 8. Export Data to CSV

After analysis, results can be written to a CSV file with `write.csv()`.

First, create a simple summary.

```{r}
analysis_result <- aggregate(
  sales ~ category,
  data = sales_data,
  FUN = sum
)

analysis_result
```

Create the output directory if necessary.

```{r}
dir.create(
  "output",
  showWarnings = FALSE
)
```

Then export the result.

```{r}
write.csv(
  analysis_result,
  "output/analysis_result.csv",
  row.names = FALSE
)
```

`row.names = FALSE` prevents R's row numbers from being written as an extra column.

The exported CSV can be opened by spreadsheet software or used by another data tool.

---

## 9. Export Text Files

`write.table()` provides more control over text output.

```{r}
write.table(
  analysis_result,
  "output/analysis_result.txt",
  sep = "\t",
  row.names = FALSE,
  quote = FALSE
)
```

Here:

- `sep = "\t"` creates tab-separated output.
- `row.names = FALSE` omits row numbers.
- `quote = FALSE` avoids adding quotation marks around character values.

---

## 10. Export Excel Files

The `writexl` package provides a simple way to create `.xlsx` files.

Install it once if necessary:

```{r eval=FALSE}
install.packages("writexl")
```

Load the package:

```{r}
library(writexl)
```

Then export a data frame.

```{r}
write_xlsx(
  analysis_result,
  "output/analysis_result.xlsx"
)
```

Excel output can be useful when analysis results need to be shared with users who work primarily with spreadsheets.

---

## 11. Save and Load R Objects

Sometimes we want to preserve an R object rather than export it as a general-purpose table.

`saveRDS()` saves one R object.

```{r}
saveRDS(
  analysis_result,
  "output/analysis_result.rds"
)
```

Load it again with `readRDS()`.

```{r}
loaded_result <- readRDS(
  "output/analysis_result.rds"
)

loaded_result
```

CSV and Excel are useful for exchanging tabular data with other tools.

RDS is useful when an R object will be saved and reused in R.

---

## 12. Build a Simple Input/Output Workflow

Now combine importing, validation, transformation, analysis, and exporting.

Load `dplyr`.

```{r}
library(dplyr)
```

Import the data and create an analysis result.

```{r}
final_result <- read.csv(
  "data/sample_sales_19.csv"
) |>
  filter(
    !is.na(category),
    !is.na(sales)
  ) |>
  group_by(category) |>
  summarise(
    total_sales = sum(sales),
    average_sales = mean(sales),
    records = n(),
    .groups = "drop"
  ) |>
  arrange(desc(total_sales))

final_result
```

Export the result.

```{r}
write.csv(
  final_result,
  "output/category_sales_summary.csv",
  row.names = FALSE
)
```

The complete workflow is:

```text
sample_sales_19.csv
       |
       v
    Import
       |
       v
    Inspect
       |
       v
   Validate
       |
       v
  Transform
       |
       v
   Analyze
       |
       v
    Export
       |
       v
category_sales_summary.csv
```

The important point is that file input and output are not isolated operations.

They are the beginning and end of a complete data analysis process.

---

## Summary

In this chapter, we learned how R communicates with external files.

| Function | Purpose |
|---|---|
| `getwd()` | Check the working directory |
| `read.csv()` | Import CSV data |
| `read.table()` | Import delimited text data |
| `read_excel()` | Import Excel data |
| `write.csv()` | Export CSV data |
| `write.table()` | Export text data |
| `write_xlsx()` | Export Excel data |
| `saveRDS()` | Save an R object |
| `readRDS()` | Load an R object |

The main idea is:

**External Data → Import → Inspect → Analyze → Export**

File paths are also part of good project design. Relative paths help make scripts portable and reproducible.

With this chapter, the Data Analysis section is complete.

In the next chapter, we will begin Practical R by connecting R directly to a database.
